---
title: "Homework-03"
format: html
editor: visual
---

```{r}
library(tidyverse)
library(janitor)
library(readxl)
library(ggeffects)
```

```{r}
salinity <- read.csv("data/salinity-pickleweed.csv")
```

## Problem 1

**a. An appropriate test**

The appropriate tests to determine the strength of the relationship between salinity and pickleweed biomass are the Pearson's correlation and Spearman rank correlation tests, telling us the significance of the effects of the predictor variable of salinity (mS/cm) on response variable pickleweed biomass (g). Pearson's correlation is to be used if there is a linear relationship between the variables, variables are continuous, variables are normally distributed, and there are independent observations. Spearman rank correlation is to be used if there is a monotonic relationship between variables and there are independent observations.

**b. Create visualization**

```{r}
salinity_preds <- ggpredict(
  salinity_lm,
  terms = "salinity_mS_cm"
)

salinity_preds
```

```{r}
ggplot(data = salinity,
       aes(x = salinity_mS_cm,
           y = pickleweed)) + 
  geom_point(fill = "tomato",
             shape = 22) + 
  geom_ribbon(data = salinity_preds,
              aes(x = x,
                  y = predicted,
                  ymin = conf.low,
                  ymax = conf.high,
                  alpha = 0.1)) + 
  geom_line(data = salinity_preds,
            aes(x = x,
                y = predicted),
            color = "blue") +
  labs(x = "Salinity (mS/cm)",
       y = "Pickleweed biomass (g)") 
```

**c. Check your assumptions and run your test**

Check assumptions

```{r}
salinity_lm <- lm(
                  pickleweed ~ salinity_mS_cm,
                  data = salinity)
```

```{r}
par(mfrow = c(2,2))
plot(salinity_lm)
```

Above, I created plots for residuals vs. fitted and scale-location to check the homoscedasticity of the observations, and because the red line is relatively straight it would suggest that the relationship between variables is linear. The Q-Q Residuals plot tells us whether there is normal distribution with the observations, and looking at this plot there is not significant deviation from the line and we can determine that there is normal distribution. Looking at the salinity data, the variables are continuous and the individual pickleweed plants are individual observations, which in conjunction with the above assumptions tells us that we can use Pearson's correlation.

Run your test

```{r}
cor.test(salinity$salinity_mS_cm,
         salinity$pickleweed,
         method = "pearson")
```

The Pearson's correlation test gives us a correlation value of 0.53.

**d. Results communication**

To evaluate the strength of the relationship between pickleweed biomass and soil salinity, I used a Pearson's correlation test and found that there is a moderate relationship (Pearson's r = 0.53, t(21) = 2.90, p = 0.0086, $\alpha$ = 0.05).

**e. Test implications**

There is a moderate relationship between salinity and pickleweed biomass, which means that there is a noticeable influence of salinity (mS/cm) on pickleweed biomass (g). Based on the results of our test, increasing salinity will lead to more success in pickleweed planting at our restoration site. However, I would note that we do not have any observations exceeding a salinity value of 10.0 mS/cm, and I don't know how surpassing this threshold would affect growth.

**f. Double check your own work**

```{r}
cor.test(salinity$salinity_mS_cm,
         salinity$pickleweed,
         method = "spearman")
```

This test would lead me to make the same decision about the null hypothesis, despite the assumptions for both test being different and this technically being the incorrect one. They give a very similar value for the strength of the relationship, which would lead me to the same conclusion in both cases.

We found a moderate relationship between salinity (mS/cm) and pickleweed biomass (g) (Spearman $\rho$ = 0.59, S = 824, p = 0.0034, $\alpha$ = 0.05).

## Problem 2. Personal data

**a. Updating your visualizations**

```{r}
personal <- read_xlsx("data/Data_Project_Water_Idea_HW3.xlsx")
```

```{r}
personal_clean <- personal |> 
  clean_names()
```


```{r}
ggplot(data = personal_clean,
       aes(x = high_daily_temperature_f,
           y = water_drank_l)) + 
  geom_jitter(height = 0,
              width = 0.3,
              color = "tomato")
```







