---
title: "Homework-03"
format: html
editor: visual
---

```{r}
library(tidyverse)
library(janitor)
library(readxl)
```

```{r}
salinity <- read.csv("data/salinity-pickleweed.csv")
```

## Problem 1

**a. An appropriate test**

The appropriate tests to determine the strength of the relationship between salinity and pickleweed biomass are the Pearson's correlation and Spearman rank correlation tests, telling us the significance of the effects of the predictor variable of salinity (mS/cm) on response variable pickleweed biomass (g). Pearson's correlation is to be used if there is a linear relationship between the variables, variables are continuous, variables are normally distributed, and there are independent observations. Spearman rank correlation is to be used if there is a monotonic relationship between variables and there are independent observations. 

**b. Create visualization**

```{r}
ggplot(data = salinity,
       aes(x = salinity_mS_cm,
           y = pickleweed)) + 
  geom_point() + 
  labs(x = "Salinity (mS/cm)",
       y = "Pickleweed biomass (g)") 
```

```{r}
salinity_lm <- lm(
                  pickleweed ~ salinity_mS_cm,
                  data = salinity)
```

```{r}
par(mfrow = c(2,2))
plot(salinity_lm)
```





